### 2024-03-12 paper/paper.qmd
updated introduction

---
title: "Evolving Dynamics: The U.S. Labor Force from 1972 to 2022 and the Enduring Challenges for Women in the Workplace"
subtitle: "My subtitle if needed"
author: 
  - Michael Fang
  - Harrison Huang
  - David James Dimalanta
thanks: "Code and data are available at: https://github.com/fanger2791/fanger2791-Enduring-Challenges-for-Women-in-the-Workplace."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

# Introduction

Over the past fifty years, the U.S. labor force has undergone significant transformations, reflecting broader socio-economic shifts, technological advancements, and changing demographic patterns. These changes have had profound implications for employment practices, workplace cultures, and the economic opportunities available to different segments of the population. However, despite these overall shifts, women in the workforce have continued to face enduring challenges, including gender pay gaps, underrepresentation in leadership positions, and the persistent juggling act of work-family balance. This paper delves into the evolving dynamics of the U.S. labor force from 1972 to 2022, with a lens on understanding how these broader changes have intersected with the experiences of women in the workplace.

The gap combines an exploration of broad labor force trends with an examination of the specific, ongoing challenges that women encounter in the workplace. While there is substantial literature on both subjects separately, there is a notable lack of comprehensive studies that integrate these perspectives to offer a detailed understanding of how historical and current labor force dynamics interact with gender-specific issues. To fill this gap, our study leverages data from the U.S. General Social Survey (GSS), to explore the nuanced ways in which the labor market's evolution over five decades has impacted women's experiences in the workforce.

The analysis of GSS data, reveals that, despite significant strides in participation rates and educational attainment among women, numerous deep-rooted barriers persist. These barriers include persistent gender segregation in certain industries and biases in hiring and promotion practices. Crucially, the survey highlights the diversity of women's experiences in the labor market, underscoring the importance of an intersectional approach to understanding workplace challenges.

The importance of this research lies in its comprehensive examination of the structural barriers that continue to limit gender equality in the workplace. By mapping the historical and contemporary landscape of the labor market with a focus on women's experiences, the paper offers valuable insights for policymakers, business leaders, and social advocates. These insights aim to inform efforts to create a more inclusive, equitable, and supportive working environment for all.

The paper is organized as follows: Section two details the methodology and source data, including the use of GSS data. Section three provides an analysis of labor force trends over the past fifty years, with an emphasis on changes in employment patterns, workforce participation, and educational attainment among women. Section four delves into the specific challenges women face in the workplace, utilizing quantitative data from the GSS and discusses the broader implications of these findings, offering targeted recommendations for policy and practice to address gender disparities.

# Data

Data used in this paper are retrieved from the US General Social Survey(GSS).
The General Social Survey (GSS) is an ongoing interview survey of U.S. households conducted by the National Opinion Research Center. It has provided a wealth of data on contemporary American society for approximately 35 years by measuring social change and trends and constants in the adult population's attitudes, behaviors, and attributes. It includes a questionnaire such as demographics, attitudes, and ideas towards certain beliefs, and even extends economic ideologies such as spending priorities. The goal is to have high-quality data available to social scientists and researchers. Over the years, the GSS has worked continuously to make the survey better. One major improvement worthy of mentioning is the addition of a Spanish version, since there is a big demographic of the U.S. that speaks Spanish, this would provide a significant increase in the coverage of U.S. household responses.

## Source Data

```{r}
#| echo: false
#| tbl-cap: Women in the workforce survey data

library(kableExtra)

variable_data <- data.frame( 
  Variable = c("fejobaff", 
               "fehire", 
               "Fefam",
               "Discaffw"), 
  Description = c("For or against preferential hiring of                       women",
                  "Should hire and promote women",
                  "Is it better for man to work, and a woman                    to tend home", 
                  "A woman won't get a job or promotion"), 
  ExampleResponse = c("Strongly Favor", 
              "Disagree", 
              "Agree",
              "Somewhat likely")
)

variable_data %>%
  kbl() %>%
  kable_classic_2(full_width = F)
```

# Results

## Respondent Demograpgics

## Respondent Labour Status

```{r}
#| label: fig-fulltime
#| fig-cap: Full Time Work Status by Gender Over Time
#| echo: false
#| warning: false
#| message: false

library(ggplot2)

data <- read.csv("../data/analysis_data/full_time_data.csv")

ggplot(data) +
  geom_line(aes(x = year, y = men, color = "Men")) +
  geom_line(aes(x = year, y = women, color = "Women")) +
  scale_color_manual(values = c("Men" = "blue", "Women" = "pink")) +
  labs(
       x = "Year",
       y = "Number of People",
       color = "Gender") +
  theme_minimal()
```

```{r}
#| label: fig-part_time
#| fig-cap: Part Time Work Status by Gender Over Time
#| echo: false
#| warning: false
#| message: false

library(ggplot2)

data <- read.csv("../data/analysis_data/part_time_data.csv")

ggplot(data) +
  geom_line(aes(x = year, y = men, color = "Men")) +
  geom_line(aes(x = year, y = women, color = "Women")) +
  scale_color_manual(values = c("Men" = "blue", "Women" = "pink")) +
  labs(
       x = "Year",
       y = "Number of People",
       color = "Gender") +
  theme_minimal()
```

```{r}
#| label: fig-parttime
#| fig-cap: Unemployed Work Status by Gender Over Time
#| echo: false
#| warning: false
#| message: false

library(ggplot2)

data <- read.csv("../data/analysis_data/unemployed_data.csv")

ggplot(data) +
  geom_line(aes(x = year, y = men, color = "Men")) +
  geom_line(aes(x = year, y = women, color = "Women")) +
  scale_color_manual(values = c("Men" = "blue", "Women" = "pink")) +
  labs(
       x = "Year",
       y = "Number of People",
       color = "Gender") +
  theme_minimal()
```
## Biases against hiring women

```{r}
#| label: fig-prefhiring
#| fig-cap: Opinions on Preferential Hiring Over Time
#| echo: false
#| warning: false
#| message: false

library(ggplot2)
library(readxl)
library(tidyr)

# Read the data from the Excel file
data <- read_excel("../data/analysis_data/pref_hiring.xlsx")

# Reshape the data to long format for easier plotting with ggplot2
data_long <- pivot_longer(data, cols = -Year, names_to = "Category", values_to = "Count")

# Plot
ggplot(data_long, aes(x = Year, y = Count, fill = Category)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal() +
  labs(
       x = "Year",
       y = "Count") +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "bottom")

```

```{r}

```

# Discussion

## Strengths of US GSS

## Potential Ethical and Bias Issues with Survey Methodology

## Potential improvements

A potential improvement that could be implemented to improve the issues mentioned above is to update the questionnaires every year. Have a psychologist screen and update the questionnaires yearly to ensure that questions are worded properly and do not initiate any thoughts from the readers. Another thing to ensure is that the language used is up to date. In addition to removing biased wording, it is also important to ensure that the words are recent enough and are not words we no longer use. Since this survey has been ongoing for a long time, such measures should maintain the quality of the survey. Lastly, the increased coverage of participants would also be very beneficial. A method could be to survey states by state ensuring that they meet a certain percentage of the population before moving on to the next state. However, this would be a very time-consuming and tedious process, not every survey has the funding and time for such an approach. These implementations would improve the quality of the survey allowing for more accurate presentations of the population.

\newpage

# References