### 2024-03-13 paper/paper.qmd
added results section to paper

---
title: "Evolving Dynamics: The U.S. Labor Force from 1972 to 2022 and the Enduring Challenges for Women in the Workplace"
subtitle: "My subtitle if needed"
author: 
  - Michael Fang
  - Harrison Huang
  - David James Dimalanta
thanks: "Code and data are available at: https://github.com/fanger2791/fanger2791-Enduring-Challenges-for-Women-in-the-Workplace."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

# Introduction

Over the past fifty years, the U.S. labor force has undergone significant transformations, reflecting broader socio-economic shifts, technological advancements, and changing demographic patterns. These changes have had profound implications for employment practices, workplace cultures, and the economic opportunities available to different segments of the population. However, despite these overall shifts, women in the workforce have continued to face enduring challenges, including gender pay gaps, underrepresentation in leadership positions, and the persistent juggling act of work-family balance. This paper delves into the evolving dynamics of the U.S. labor force from 1972 to 2022, with a lens on understanding how these broader changes have intersected with the experiences of women in the workplace.

The gap combines an exploration of broad labor force trends with an examination of the specific, ongoing challenges that women encounter in the workplace. While there is substantial literature on both subjects separately, there is a notable lack of comprehensive studies that integrate these perspectives to offer a detailed understanding of how historical and current labor force dynamics interact with gender-specific issues. To fill this gap, our study leverages data from the U.S. General Social Survey (GSS), to explore the nuanced ways in which the labor market's evolution over five decades has impacted women's experiences in the workforce.

The analysis of GSS data, reveals that, despite significant strides in participation rates and educational attainment among women, numerous deep-rooted barriers persist. These barriers include persistent gender segregation in certain industries and biases in hiring and promotion practices. Crucially, the survey highlights the diversity of women's experiences in the labor market, underscoring the importance of an intersectional approach to understanding workplace challenges.

The importance of this research lies in its comprehensive examination of the structural barriers that continue to limit gender equality in the workplace. By mapping the historical and contemporary landscape of the labor market with a focus on women's experiences, the paper offers valuable insights for policymakers, business leaders, and social advocates. These insights aim to inform efforts to create a more inclusive, equitable, and supportive working environment for all.

The paper is organized as follows: Section two details the methodology and source data, including the use of GSS data. Section three provides an analysis of labor force trends over the past fifty years, with an emphasis on changes in employment patterns, workforce participation, and educational attainment among women. Section four delves into the specific challenges women face in the workplace, utilizing quantitative data from the GSS and discusses the broader implications of these findings, offering targeted recommendations for policy and practice to address gender disparities.

# Data

Data used in this paper are retrieved from the US General Social Survey(GSS).
The General Social Survey (GSS) is an ongoing interview survey of U.S. households conducted by the National Opinion Research Center. It has provided a wealth of data on contemporary American society for approximately 35 years by measuring social change and trends and constants in the adult population's attitudes, behaviors, and attributes. It includes a questionnaire such as demographics, attitudes, and ideas towards certain beliefs, and even extends economic ideologies such as spending priorities. The goal is to have high-quality data available to social scientists and researchers. Over the years, the GSS has worked continuously to make the survey better. One major improvement worthy of mentioning is the addition of a Spanish version, since there is a big demographic of the U.S. that speaks Spanish, this would provide a significant increase in the coverage of U.S. household responses.

## Source Data

```{r}
#| echo: false
#| tbl-cap: Women in the workforce survey data

library(kableExtra)

variable_data <- data.frame( 
  Variable = c("fejobaff", 
               "fehire", 
               "Fefam",
               "Discaffw"), 
  Description = c("For or against preferential hiring of                       women",
                  "Should hire and promote women",
                  "Is it better for man to work, and a woman                    to tend home", 
                  "A woman won't get a job or promotion"), 
  ExampleResponse = c("Strongly Favor", 
              "Disagree", 
              "Agree",
              "Somewhat likely")
)

variable_data |>
  kbl() |>
  kable_classic_2(full_width = F)
```

# Results

## Respondent Demograpgics

## Respondent Labour Status

```{r}
#| label: fig-fulltime
#| fig-cap: Full Time Work Status by Gender Over Time
#| echo: false
#| warning: false
#| message: false

library(ggplot2)

data <- read.csv("../data/analysis_data/full_time_data.csv")

ggplot(data) +
  geom_line(aes(x = year, y = men, color = "Men")) +
  geom_line(aes(x = year, y = women, color = "Women")) +
  scale_color_manual(values = c("Men" = "blue", "Women" = "pink")) +
  labs(
       x = "Year",
       y = "Number of People",
       color = "Gender") +
  theme_minimal()
```

```{r}
#| label: fig-part_time
#| fig-cap: Part Time Work Status by Gender Over Time
#| echo: false
#| warning: false
#| message: false

library(ggplot2)

data <- read.csv("../data/analysis_data/part_time_data.csv")

ggplot(data) +
  geom_line(aes(x = year, y = men, color = "Men")) +
  geom_line(aes(x = year, y = women, color = "Women")) +
  scale_color_manual(values = c("Men" = "blue", "Women" = "pink")) +
  labs(
       x = "Year",
       y = "Number of People",
       color = "Gender") +
  theme_minimal()
```

```{r}
#| label: fig-parttime
#| fig-cap: Unemployed Work Status by Gender Over Time
#| echo: false
#| warning: false
#| message: false

library(ggplot2)

data <- read.csv("../data/analysis_data/unemployed_data.csv")

ggplot(data) +
  geom_line(aes(x = year, y = men, color = "Men")) +
  geom_line(aes(x = year, y = women, color = "Women")) +
  scale_color_manual(values = c("Men" = "blue", "Women" = "pink")) +
  labs(
       x = "Year",
       y = "Number of People",
       color = "Gender") +
  theme_minimal()
```
## Biases against hiring women

```{r}
#| label: fig-prefhiring
#| fig-cap: Opinions on Preferential Hiring Over Time
#| echo: false
#| warning: false
#| message: false

library(ggplot2)
library(readxl)
library(tidyr)

# Read the data from the Excel file
data <- read_excel("../data/analysis_data/pref_hiring.xlsx")

# Reshape the data to long format for easier plotting with ggplot2
data_long <- pivot_longer(data, cols = -Year, names_to = "Category", values_to = "Count")

# Plot
ggplot(data_long, aes(x = Year, y = Count, fill = Category)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal() +
  labs(
       x = "Year",
       y = "Count") +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "bottom")

```

# Results
```{r}
#| label: fig-fehire
#| fig-cap: Attitudes Towards Hiring and Promoting Women
#| echo: false
#| warning: false
#| message: false

library(ggplot2)
library(readxl)
library(tidyr)

# read in data
fehire_data <- read_csv("../data/analysis_data/fehire_data.csv")

# Format year data as numeric for line graphs
fehire_data$year <- as.numeric(fehire_data$year)

# Reshape the data to long format for line graphs and convert 'count' to numeric
fehire_data_long <- fehire_data |>
  pivot_longer(cols = c(strongly_agree, agree, disagree, strongly_disagree),
               names_to = "Response",
               values_to = "Count") |>
  mutate(Count = as.numeric(Count))

# fehire line graph
ggplot(fehire_data_long, aes(x = year, y = Count, group = Response, color = Response)) +
  geom_line() +
  scale_y_continuous(limits = c(0, 1000)) +
  labs(title = "Fehire Responses Over Time",
       x = "Year",
       y = "Count",
       color = "Response Category") +
  theme_minimal()
```

```{r}
#| label: fig-fefam
#| fig-cap: Views on Gender Roles in Work and Family Life
#| echo: false
#| warning: false
#| message: false

library(ggplot2)
library(readxl)
library(tidyr)

# read in data
fefam_data <- read_csv("../data/analysis_data/fefam_data.csv")

# Format year data as numeric for line graphs
fefam_data$year <- as.numeric(fefam_data$year)

# Reshape the data to long format for line graphs and convert 'count' to numeric
fefam_data_long <- fefam_data |>
  pivot_longer(cols = c(strongly_agree, agree, disagree, strongly_disagree),
               names_to = "Response to: better for man to work, woman tend home",
               values_to = "count") |>
  mutate(count = as.numeric(count))

# fefam line graph
ggplot(fefam_data_long, aes(x = year, y = count, group = `Response to: better for man to work, woman tend home`, color = `Response to: better for man to work, woman tend home`)) +
  geom_line() +
  scale_y_continuous(limits = c(0, 1200)) +
  labs(title = "Responses to: better for man to work, woman tend home",
       x = "Year",
       y = "Count") +
  theme_minimal()
```

```{r}
#| label: fig-discaffw
#| fig-cap: Perceptions of Discrimination Against Women in Job Promotions
#| echo: false
#| warning: false
#| message: false

library(ggplot2)
library(readxl)
library(tidyr)

# read in data
discaffw_data <- read_csv("../data/analysis_data/discaffw_data.csv")

# Format year data as numeric for line graphs
discaffw_data$year <- as.numeric(discaffw_data$year)

# Reshape the data to long format for line graphs and convert 'count' to numeric
discaffw_data_long <- discaffw_data |>
  pivot_longer(cols = c(very_likely, somewhat_likely, somewhat_unlikely, very_unlikely),
               names_to = "Response",
               values_to = "Count") |>
  mutate(Count = as.numeric(Count))

# discaffw line graph
ggplot(discaffw_data_long, aes(x = year, y = Count, group = Response, color = Response)) +
  geom_line() +
  scale_y_continuous(limits = c(0, 700)) +
  labs(title = "Discaffw Responses Over Time",
       x = "Year",
       y = "Count",
       color = "Response Category") +
  theme_minimal()
```

```{r}
#| label: fig-education
#| fig-cap: Educational Attainment by Gender
#| echo: false
#| warning: false
#| message: false

library(ggplot2)
library(readxl)
library(tidyr)

# read in data
education_data <- read_csv("../data/analysis_data/education_data.csv")

# Format year data as numeric for line graphs
education_data$year <- as.numeric(education_data$year)

# Reshape the data to long format for line graphs and convert 'count' to numeric
education_data_long <- education_data |>
  pivot_longer(cols = c(less_than_highschool, high_school, junior_college, bachelor, graduate),
               names_to = "Education Level",
               values_to = "Count") |>
  mutate(Count = as.numeric(Count))

# education line graph
ggplot(education_data_long, aes(x = year, y = Count, group = `Education Level`, color = `Education Level`)) +
  geom_line() +
  labs(title = "Education Levels Over Time",
       x = "Year",
       y = "Count",
       color = "Education Level") +
  theme_minimal()
```

```{r}
#| label: fig-eduTojob
#| fig-cap: Trends in Full-Time Work and Educational Attainment for Women
#| echo: false
#| warning: false
#| message: false

library(ggplot2)
library(readxl)
library(tidyr)

# read in data
combined_data <- read_csv("../data/analysis_data/combined_data.csv")

# combined education and full-time data in line graph
ggplot(combined_data, aes(x = as.numeric(as.character(year)), y = Count, color = category, group = category)) +
  geom_line() +
  scale_y_continuous(name = "Count", limits = c(0, 1000)) +
  labs(title = "Comparison of Full-Time Work for Women with Higher Education Levels",
       x = "Year",
       y = "Count",
       color = "Category") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Attitudes Towards Hiring and Promoting Women
@fig-fehire describes the GSS data on attitudes towards hiring and promoting women ("FEHIRE") from 1996 to 2022, showing a progressive shift towards more positive attitudes. The proportion of respondents strongly agreeing with the need to hire and promote women has seen a significant increase, particularly noticeable in the year 2021, where 620 respondents strongly agreed, compared to 244 in 1996. This trend suggests a growing consensus on the importance of gender equality in the workplace over the analyzed period. Despite this positive shift, there remains a consistent portion of respondents who disagree to varying degrees, highlighting that there is still work to be done to achieve universal acceptance of this principle.

## Perceptions of Discrimination Against Women in Job Promotions
@fig-discaffw describes data on perceptions of discrimination against women in job promotions ("DISCAFFW") from 1996 to 2022, also indicating evolving public opinion. While the early years showed a more divided view on the likelihood of a woman not getting a job or promotion due to discrimination, there is a noticeable decline in the percentage of respondents who find it "very likely" or "somewhat likely." However, the absence of data in 2021 for this variable suggests a gap in understanding public opinion for that year. The overall trend suggests an increasing awareness and potentially decreasing concern over direct discrimination as a barrier to women's career advancement, though substantial concern remains.

## Views on Gender Roles in Work and Family Life
@fig-fefam describes the survey responses to "FEFAM," concerning the belief that it's better for men to work and women to tend to the home, showing significant changes from 1977 to 2022. Early in the survey period, a substantial number of respondents agreed with traditional gender roles. Over time, there's been a notable shift towards disagreement with this view, especially strong in the 2021 data, where 1,146 strongly disagreed. This evolution reflects broader societal changes towards gender roles within family and work life, emphasizing a move towards more egalitarian views.

## Educational Attainment by Gender
@fig-education shows the GSS data on educational attainment, segmented by gender from 1972 to 2022, revealing trends in the pursuit and achievement of higher education. Both males and females have shown increased attainment of bachelor's and graduate degrees over time. Notably, the gap in educational attainment between males and females has narrowed, with women showing significant gains in both bachelor's and graduate degrees, especially highlighted in the 2021 data. This trend underscores the increasing value placed on education for all individuals, regardless of gender, and reflects the broader societal shifts towards gender equality in education.

## Full-Time Work for Women
@fig-eduTojob reveals a significant upward trend in the number of women engaged in full-time employment over the years. Starting from 191 women in 1972, the numbers generally increased, peaking at 963 women in full-time work in 2006, before experiencing fluctuations and settling at 728 in 2022. This progression underscores the growing integration of women into the workforce and the breaking down of traditional gender roles that confined women to domestic responsibilities. The notable peak in 2006 could be indicative of economic, social, and legislative changes that encouraged or necessitated higher female participation in the workforce.

## Educational Attainment by Gender
Parallel to the rise in full-time employment, @fig-eduTojob also highlights significant strides in women's educational attainment. The data tracks the number of women who have achieved bachelor's and graduate degrees from 1972 through 2022. In 1972, 45 women held bachelor's degrees and 10 held graduate degrees. By 2022, these numbers had surged to 353 for bachelor's degrees and 228 for graduate degrees, reflecting a substantial increase in higher education attainment among women.

This trend not only demonstrates the shifting societal expectations and the dismantling of barriers to women's education but also aligns with the increased demand for higher education in a knowledge-driven economy. The remarkable increase in graduate degrees highlights the progress toward gender parity in advanced education levels, potentially contributing to narrowing the gender gap in professional and academic fields.

# Discussion

## Strengths of US GSS

Interviews have been conducted with a total of 51,020 respondents. The 1972-74 surveys used modified probability designs and the remaining surveys were completed using a
full-probability sample design, producing a high-quality, representative sample of the adult population of the U.S. The GSS has a response rate of over 70 percent above that of other major social science surveys and 40-45 percentage points higher than the industry average.

## Potential Ethical and Bias Issues with Survey Methodology

In the Fehelp variable, the question is framed as “It is more important for a wife to help her husband's career than to have one herself” and then participants are asked to respond from Strongly agree, agree, disagree, to disagree strongly. The main issue with these types of questions is the wording and the answer scales. This wordy question, could lose the reader or create misunderstanding. In a lengthy survey, it is very important to keep questions concise and straightforward. The question could be rephrased as: “Whose career is more important in a relationship?” and then the scale of wife and husband on a scale. This way it would not prompt the participant to think in any direction before actually answering the question thus reducing biases that might have occurred otherwise. Another thing worthy to mention is the percentage of participants in the actual population. Simply looking at the number of total respondents may seem very impressive, however when compared on a large scale, it is less than 0.0001% of the population. Since the percentage is so low, it is very difficult to make generalizations about the actual population like this.

## Potential improvements

A potential improvement that could be implemented to improve the issues mentioned above is to update the questionnaires every year. Have a psychologist screen and update the questionnaires yearly to ensure that questions are worded properly and do not initiate any thoughts from the readers. Another thing to ensure is that the language used is up to date. In addition to removing biased wording, it is also important to ensure that the words are recent enough and are not words we no longer use. Since this survey has been ongoing for a long time, such measures should maintain the quality of the survey. Lastly, the increased coverage of participants would also be very beneficial. A method could be to survey states by state ensuring that they meet a certain percentage of the population before moving on to the next state. However, this would be a very time-consuming and tedious process, not every survey has the funding and time for such an approach. These implementations would improve the quality of the survey allowing for more accurate presentations of the population.

\newpage

# References